### 2023-03-15 rebuttal_response.tex

\documentclass[12pt]{article}
%\usepackage[utf8]{inputenc}

\usepackage[margin=0.97in]{geometry}
\usepackage{graphicx}
% DO NOT USE \usepackage{times}, it will be removed by typesetters
%\usepackage{times}

\usepackage{comment}
\usepackage{amsmath,amssymb} % define this before the line numbering.
\usepackage{color}

% The "axessiblity" package can be found at: https://ctan.org/pkg/axessibility?lang=en
\usepackage[accsupp]{axessibility}  % Improves PDF readability for those with disabilities.

%============================================================================
% Include other packages here
%============================================================================
\usepackage{graphicx}
%\usepackage{amsthm}           % for proof
\usepackage{booktabs}

\usepackage{epsfig}
\usepackage{tabularx}
\usepackage[table]{xcolor}       % for colored stuff
% \usepackage{verbatim}            % For multiline comments
\usepackage{booktabs}
\usepackage{relsize}
\usepackage{multirow}
\usepackage{scrextend}           % Put footmark in the bottom of the page
\usepackage{array}               % Thicker vertical rules in tables, lines in only certain columns
\usepackage{makecell}            % For thicker horizontal rules in tables
\usepackage{subcaption}          % For subfigures
\usepackage{wasysym}             % For right filled arrow
\usepackage{lipsum}              % For random filler text
\usepackage{esvect}              % For vv command arrow above symbols
\usepackage{soul}                % For striked text command \st and hightlight
\usepackage{float}
\usepackage{caption}
\usepackage{bm}
\usepackage{cite}                % orders the citations in increasing order
\usepackage{paralist}            % for compactitem
\usepackage{enumitem}            % for noindent itemize
\usepackage{empheq}
\usepackage{gensymb}             % for degree symbol
\usepackage{wrapfig}             % for wraptable
\usepackage{xspace}              % for dynamic controlled space at end of abbreviation

\definecolor{lightgreen}{HTML}{90EE90}
\newcommand{\coloredeq}[2]{\begin{empheq}[box=\colorbox{lightgreen}]{align}\label{#1}#2\end{empheq}}
\newcommand{\NParagraph}[1]{\vspace{0.6mm} \noindent \textbf{#1}
\hspace{0mm}}

\captionsetup{font=small}        % Figures and Tables would have small captions 
\allowdisplaybreaks              % Break aligned equations

\graphicspath{{images/}}

%===============================================================================
% For diagrams in Latex
%===============================================================================
% https://tex.stackexchange.com/a/398269
\makeatletter
\@namedef{ver@everyshi.sty}{}
\makeatother
\usepackage{tikz} % for some figures
\usetikzlibrary{calc}  % for centerarc
\usetikzlibrary{arrows}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{shapes.arrows}
\usetikzlibrary{fadings, shadows}
\usetikzlibrary{decorations.text}
\def\centerarc[#1](#2)(#3:#4:#5)% Syntax: [draw options] (center) (initial angle:final angle:radius)
    { \draw[#1] ($(#2)+({#5*cos(#3)},{#5*sin(#3)})$) arc (#3:#4:#5); }

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}

\input{notations}

\newcommand{\revOne}{{\color{red}{R1}}}
\newcommand{\revTwo}{{\color{orange}{R2}}}
\newcommand{\revThree}{{\color{my_magenta}{R3}}}
\newcommand{\revFour}{{\color{my_blue}{R4}}}

\usepackage[pagebackref,breaklinks,colorlinks,citecolor={my_blue},urlcolor=black]{hyperref}
% Support for easy cross-referencing
\usepackage[capitalize]{cleveref}
\crefname{section}{Sec.}{Secs.}
\Crefname{section}{Section}{Sections}
\Crefname{table}{Table}{Tables}
\crefname{table}{Tab.}{Tabs.}

%\title{ Simultaneous Localization and Mapping for Dynamic Scenes}
%\author{Xiaoming Liu}
\date{} % clear date

%============================================================================
%============================================================================
%============================================================================
\begin{document}

\pagenumbering{arabic}% resets `page` counter to 1
\renewcommand*{\thepage}{Page \arabic{page}}
%============================================================================
%============================================================================
%============================================================================

        \begin{table}[!tb]
            \caption{\textbf{Datasets} and their usage. [Key: Det.= Detection, Illu.= Illumination, Comp. Ego Movement= Ego Movement, *= Do not exist]}
            \label{tab:datasets}
            \centering
            \scalebox{\scaleFraction}
            {
            % \footnotesize
            % \setlength\tabcolsep{0.15cm}
            \begin{tabular}{m l t cc t cc t cccc t c c|cm}
                \myTopRule
                \multirow{2}{*}{\textbf{Dataset} {\raisebox{0.1\normalbaselineskip}{\scalebox{0.85}{\downarrowRHD}}} } & \multirow{2}{*}{Train} & \multirow{2}{*}{Test} & \multicolumn{2}{ct}{Task} & \multicolumn{4}{ct}{Weather} & 
                \multirow{2}{*}{Night} & {Comp. Ego}\\
                \cline{4-9}
                &&& Det. & SLAM & Rain & Dust & Fog & Snow && Movement\\
                \myTopRule
                \kitti \cite{geiger2012we}                & \cmark    & \cmark & \cmark    & \cmark    & \mathDash & \mathDash & \mathDash & \mathDash & \mathDash & \mathDash\\
                \waymo \cite{sun2020scalability}          & \cmark    & \cmark & \cmark    & \cmark    & \cmark    & \cmark    & \cmark    & \mathDash & \cmark    & \mathDash  \\
                \nuscenes \cite{caesar2020nuscenes}       & \mathDash & \cmark & \cmark    & \cmark    & \cmark    & \mathDash & \mathDash & \mathDash & \cmark    & \mathDash \\
                Adv Weather\cite{bijelic2020seeing}       & \mathDash & \cmark & \cmark    & \mathDash & \cmark    & \mathDash & \cmark    & \cmark    & \cmark    & \mathDash \\
                \ithaca\cite{diaz2022ithaca365}           & \mathDash & \cmark & \cmark    & \cmark    & \cmark    & \mathDash & \cmark    & \cmark    & \cmark    & \mathDash\\
                \argoverseTwo  \cite{wilson2023argoverse} & \cmark    & \cmark & \mathDash & \cmark    & \mathDash & \mathDash & \mathDash & \mathDash & \cmark & \mathDash \\
                \rugd \cite{weigness2019rugd}             & \mathDash & \cmark & \mathDash & \cmark    & \multicolumn{4}{c|}{\text{Unconstrained Off-Road}} & \cmark    & \cmark\\
                MSU Canvas*                               & \mathDash & \cmark & \mathDash & \cmark    & \cmark    & \cmark    & \cmark    & \cmark    & \cmark    & \cmark\\
                Shaky \kitti *                            & \mathDash & \cmark & \mathDash & \cmark    & \cmark    & \cmark    & \cmark    & \cmark    & \cmark    & \cmark\\
                \myTopRule
            \end{tabular}
            }
        \end{table}

    Great point! We plan to evaluate our system from the following perspectives:

%============================================================================
%============================================================================
%============================================================================
{\small
\bibliographystyle{ieee_fullname}
\bibliography{references}
}

\end{document}